---
title: "Handler Effect"
author: "Michelle Bang"
date: '2022-11-04'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(dplyr)
```

```{r}
data <- read.table(file = "TransmissionDataExtended_forMichelle_Set1_Male_Oct22.xlsx - All_male_data_18192021.tsv",
                   header = TRUE,
                   sep = "\t")
```

```{r}
data$percentage <- data$number_of_GFP_kernels / (data$number_of_GFP_kernels + data$number_of_WT_kernels)
data$totalkernels <- data$number_of_GFP_kernels + data$number_of_WT_kernels
```

```{r}
data <- data %>%
  mutate(pollen_handler = if_else(pollen_handler == 0, "?", pollen_handler))
```


```{r}
data %>%
  group_by(pollen_handler) %>%
  summarize(avg_per = mean(percentage), count = n())
```
```{r}
data %>%
  group_by(pollen_handler, year) %>%
  summarize(avg_per = mean(percentage), count = n())
```
```{r}
data %>%
  filter(year == 2018) %>%
  group_by(pollen_handler) %>%
  summarize(avg_per = mean(percentage), count = n())
```
```{r}
data %>%
  filter(year == 2019) %>%
  group_by(pollen_handler) %>%
  summarize(avg_per = mean(percentage), count = n())
```

```{r}
data %>%
  filter(year == 2020) %>%
  group_by(pollen_handler) %>%
  summarize(avg_per = mean(percentage), count = n())
```

```{r}
data %>%
  filter(year == 2021) %>%
  group_by(pollen_handler) %>%
  summarize(avg_per = mean(percentage), count = n())
```

```{r}
data %>%
  group_by(allele) %>%
  summarize(count = n()) %>%
  arrange(desc(count))
```
```{r}
data %>%
  group_by(allele, pollen_handler) %>%
  summarize(count = n()) %>%
  arrange(desc(count))
```


```{r}
# has only one handler
data %>%
  group_by(allele, pollen_handler) %>%
  summarize(count = n()) %>%
  arrange(desc(count)) %>%
  filter(allele == "R106E07")
```
```{r}
ggplot(data = data, mapping = aes(x = allele, y = percentage, color = pollen_handler)) +
  geom_point()
```

```{r}
ggplot(data = data, mapping = aes(x = pollen_handler, y = percentage, color = allele)) +
  geom_point()
```

```{r}
data %>%
  group_by(allele) %>%
  summarize(mean_kernel = mean(totalkernels)) %>%
  arrange(desc(mean_kernel))
```
```{r}
data %>%
  group_by(allele) %>%
  summarize(mean_wtkernel = mean(number_of_WT_kernels)) %>%
  arrange(desc(mean_wtkernel))
```

```{r}
data %>%
  group_by(allele) %>%
  summarize(mean_gfpkernel = mean(number_of_GFP_kernels)) %>%
  arrange(desc(mean_gfpkernel))
```

---R84A12---

```{r}
data %>%
  group_by(allele, pollen_handler) %>%
  summarize(count = n(),meankernel = mean(totalkernels), sdkernel = sd(totalkernels), meanp = mean(percentage)) %>%
  arrange(desc(count)) %>%
  filter(allele == "R84A12")
# mainly focus on the top two 
# for actual statistics, focus on top two (? and J)
# for visualization, try using all
# special allele
# also see if total kernels are different over years and handlers 
# then see if difference in percentage comes from total counts
# and have response: percentage, explanatory: handler, year, totalcount
# add plots at the end as well
```

```{r}
# visualization of R84A12
R84A12 <- data %>%
  filter(allele == "R84A12")

ggplot(R84A12, mapping = aes(x = totalkernels, y = percentage, color = factor(year))) +
  geom_point()
```

```{r}
# linear model for one allele (R84A12)
# had handler ?, H, J, M, Z
data2 <- data %>%
  filter(allele =="R84A12")
modR84A12 <- lm(percentage ~ pollen_handler + factor(year), data2)
summary(modR84A12)
# would have to do anova with reduced model to see p-value 

data2 <- data %>%
  filter(allele =="R84A12")
mod2R84A12 <- lm(percentage ~ pollen_handler, data2)
summary(mod2R84A12)


```

```{r}
anova(mod2R84A12, modR84A12)
# p-value of .01565
```

```{r}
# linear model for one allele (R84A12)
# looking at totalkernels (think about doing a log transform on this model)
# look at AIC to see which one works better 

mod3R84A12 <- lm(percentage ~ totalkernels, R84A12)
summary(mod3R84A12)
#more competition.. less difference in percentage..
```

```{r}
#find other 2 handlers to compare except ?
redR84A12 <- R84A12 %>% 
  filter(pollen_handler == "J" | pollen_handler == "?")
t.test(percentage ~ pollen_handler, data = redR84A12)
```

```{r}
mod4R84A12 <- lm(totalkernels ~ factor(year) + pollen_handler + female_genotype_group, redR84A12)
summary(mod4R84A12)

lmod4R84A12 <- lm(log(totalkernels) ~ factor(year) + pollen_handler + female_genotype_group, redR84A12)
summary(lmod4R84A12)

```

```{r}
mod5R84A12 <- lm(percentage ~ factor(year) + pollen_handler + totalkernels, redR84A12)
summary(mod5R84A12)
mod6R84A12 <- lm(percentage ~ factor(year) + pollen_handler + totalkernels, R84A12)
summary(mod6R84A12)
```

---R96B12---

```{r}
# J, Z and ?
#top 2 is J and Z

data %>%
  group_by(allele, pollen_handler) %>%
  summarize(count = n(), meankernel = mean(totalkernels), sdkernel = sd(totalkernels), meanp = mean(percentage)) %>%
  arrange(desc(count)) %>%
  filter(allele == "R96B12")
```

```{r}
# visualization of R96B12
R96B12 <- data %>%
  filter(allele =="R96B12")

ggplot(R96B12, mapping = aes(x = totalkernels, y = percentage, color = pollen_handler)) +
  geom_point()
```

```{r}
modR96B12 <- lm(percentage ~ pollen_handler + factor(year), R96B12)
summary(modR96B12)

mod2R96B12 <- lm(percentage ~ pollen_handler, R96B12)
summary(mod2R96B12)
```

```{r}
anova(mod2R96B12, modR96B12)
# p-value of .2101
```

```{r}
#R96B12
# J and Z only
# very low r-square values
# not much difference in handlers
# next step: total kernel effect
redR96B12 <- data %>%
  filter(allele =="R96B12") %>%
  filter(pollen_handler != "?")
mod2HR96B12 <- lm(percentage ~ pollen_handler, redR96B12)
summary(mod2HR96B12)
```

```{r}
mod3R96B12 <- lm(percentage ~ totalkernels, R96B12)
summary(mod3R96B12)
```

```{r}
t.test(percentage ~ pollen_handler, data = redR96B12)
```

```{r}
mod4R96B12 <- lm(totalkernels ~ factor(year) + pollen_handler + female_genotype_group, redR96B12)
summary(mod4R96B12)

lmod4R96B12 <- lm(log(totalkernels) ~ factor(year) + pollen_handler + female_genotype_group, redR96B12)
summary(lmod4R96B12)

```

```{r}
mod5R96B12 <- lm(percentage ~ factor(year) + pollen_handler + totalkernels, redR96B12)
summary(mod5R96B12)

mod6R96B12 <- lm(percentage ~ factor(year) + pollen_handler + totalkernels, R96B12)
summary(mod6R96B12)
```

---R96C12---

```{r}
data %>%
  group_by(allele, pollen_handler, year)%>%
  summarize(count = n(), meankernel = mean(totalkernels), sdkernel = sd(totalkernels), meanp = mean(percentage)) %>%
  arrange(desc(count)) %>%
  filter(allele == "R96C12")
# possible to do regression on this... (two sample t-test works)
# ^ not putting the year in
# focus is on consistency
# makes sense to compare M and J (t-test)
# this allele is likely to have a handler effect 
# good to focus on this allele
# for each row, percentage and s.d. and total number of kernels (ex. 10 data points for C)
# x-axis is each row, and total kernels on y-axis
# possion regression
# or take the number of kernels as the response and do linear regression (lm... do log transformations maybe) and explanatory variable is handler and year (probably won't need interaction term)
# for mean kernel, look into 3 genotypes (does depend on the years)... include in lm
```

```{r}
# visualization of R96C12
R96C12 <- data %>%
  filter(allele =="R96C12")

ggplot(R96C12, mapping = aes(x = totalkernels, y = percentage, color = pollen_handler)) +
  geom_point()
```

```{r}
modR96C12 <- lm(percentage ~ pollen_handler + factor(year), R96C12)
summary(modR96C12)

mod2R96C12 <- lm(percentage ~ pollen_handler, R96C12)
summary(mod2R96C12)
```

```{r}
anova(mod2R96C12, modR96C12)
# p-value of .01406
```

```{r}
#R96C12... want to compare M and J
redR96C12 <- data %>%
  filter(allele =="R96C12") %>%
  filter(pollen_handler == "M" | pollen_handler == "J")
mod2HR96C12 <- lm(percentage ~ pollen_handler, redR96C12)
summary(mod2HR96C12)
```

```{r}
mod3R96C12 <- lm(percentage ~ totalkernels, R96C12)
summary(mod3R96C12)
```

```{r}
t.test(percentage ~ pollen_handler, data = redR96C12)
```

```{r}
mod4R96C12 <- lm(totalkernels ~ pollen_handler, redR96C12)
summary(mod4R96C12)

lmod4R96C12 <- lm(log(totalkernels) ~ pollen_handler, redR96C12)
summary(lmod4R96C12)

```

```{r}
mod5R96C12 <- lm(percentage ~ pollen_handler + totalkernels, redR96C12)
summary(mod5R96C12)

#take this to all handlers
mod6R96C12 <- lm(percentage ~ pollen_handler + totalkernels, R96C12)
summary(mod6R96C12)
```

---R102H01---
pollen handler might have an effect on total kernels. year, pollen handler, and total handler has no effect on percentage. 

```{r}
data %>%
  group_by(allele, pollen_handler)%>%
  summarize(count = n(), meankernel = mean(totalkernels), sdkernel = sd(totalkernels), meanp = mean(percentage)) %>%
  arrange(desc(count)) %>%
  filter(allele == "R102H01")
```

```{r}
# visualization of R102H01
R102H01 <- data %>%
  filter(allele =="R102H01")

ggplot(R102H01, mapping = aes(x = totalkernels, y = percentage, color = pollen_handler)) +
  geom_point()
```

```{r}
modR102H01 <- lm(percentage ~ pollen_handler + factor(year), R102H01)
summary(modR102H01)

mod2R102H01 <- lm(percentage ~ pollen_handler, R102H01)
summary(mod2R102H01)
```

```{r}
anova(mod2R102H01, modR102H01)
# p-value of .3057
# year is not significant
```

```{r}
#R102H01... want to compare Z and J
#no difference
redR102H01 <- data %>%
  filter(allele =="R102H01") %>%
  filter(pollen_handler == "Z" | pollen_handler == "J")
mod2HR102H01 <- lm(percentage ~ pollen_handler, redR102H01)
summary(mod2HR102H01)
```

```{r}
#p-value .0929
mod3R102H01 <- lm(percentage ~ totalkernels, R102H01)
summary(mod3R102H01)
```

```{r}
#no diff
t.test(percentage ~ pollen_handler, data = redR102H01)
```

```{r}
mod4R102H01 <- lm(totalkernels ~ pollen_handler, redR102H01)
summary(mod4R102H01)
#p-value: .0268

lmod4R102H01 <- lm(log(totalkernels) ~ pollen_handler, redR102H01)
summary(lmod4R102H01)
#p-value: .0277
```

```{r}
mod5R102H01 <- lm(percentage ~ pollen_handler + totalkernels, redR102H01)
summary(mod5R102H01)

#take this to all handlers
mod6R102H01 <- lm(percentage ~ pollen_handler + totalkernels, R102H01)
summary(mod6R102H01)
```

---R49F11---
nothing of significance

```{r}
data %>%
  group_by(allele, pollen_handler)%>%
  summarize(count = n(), meankernel = mean(totalkernels), sdkernel = sd(totalkernels), meanp = mean(percentage)) %>%
  arrange(desc(count)) %>%
  filter(allele == "R49F11")
```

```{r}
# visualization of R49F11
R49F11 <- data %>%
  filter(allele =="R49F11")

ggplot(R49F11, mapping = aes(x = totalkernels, y = percentage, color = pollen_handler)) +
  geom_point()
```

```{r}
modR49F11 <- lm(percentage ~ pollen_handler + factor(year), R49F11)
summary(modR49F11)

mod2R49F11 <- lm(percentage ~ pollen_handler, R49F11)
summary(mod2R49F11)
```

```{r}
# no year effect
anova(mod2R49F11, modR49F11)
# p-value of .6477
```

## decided not compare two handlers for this one

```{r}
# not significant
mod3R49F11 <- lm(percentage ~ totalkernels, R49F11)
summary(mod3R49F11)
```


```{r}
# nothing is significant.
lmod4R49F11 <- lm(log(totalkernels) ~ pollen_handler, R49F11)
summary(lmod4R49F11)

```

```{r}
# not significant
mod6R49F11<- lm(percentage ~ pollen_handler + totalkernels, R49F11)
summary(mod6R49F11)
```
---R82A03---
no significance

```{r}
data %>%
  group_by(allele, pollen_handler)%>%
  summarize(count = n(), meankernel = mean(totalkernels), sdkernel = sd(totalkernels), meanp = mean(percentage)) %>%
  arrange(desc(count)) %>%
  filter(allele == "R82A03")
```

```{r}
# visualization of R82A03
R82A03<- data %>%
  filter(allele =="R82A03")

ggplot(R82A03, mapping = aes(x = totalkernels, y = percentage, color = pollen_handler)) +
  geom_point()
```

```{r}
modR82A03 <- lm(percentage ~ pollen_handler + factor(year), R82A03)
summary(modR82A03)

mod2R82A03<- lm(percentage ~ pollen_handler, R82A03)
summary(mod2R82A03)
```

```{r}
# year not significant
anova(mod2R82A03, modR82A03)
# p-value of .7363
```
## decided not to compare two handlers

```{r}
# not signficant
mod3R82A03 <- lm(percentage ~ totalkernels, R82A03)
summary(mod3R82A03)
```

```{r}
# not signficant
lmod4R82A03 <- lm(log(totalkernels) ~ pollen_handler, R82A03)
summary(lmod4R82A03)
```

```{r}
# not signficant
mod6R82A03 <- lm(percentage ~ pollen_handler + totalkernels, R82A03)
summary(mod6R82A03)
```
---R89B08---
more kernels, lower the percentage...
2019 vs. 2020 there is a diff (J)

```{r}
# could be interesting to compare J for 2019 vs. 2020
data %>%
  group_by(allele, pollen_handler, year)%>%
  summarize(count = n(), meankernel = mean(totalkernels), sdkernel = sd(totalkernels), meanp = mean(percentage)) %>%
  arrange(desc(count)) %>%
  filter(allele == "R89B08")
```

```{r}
# visualization of R89B08
# negative slope
R89B08 <- data %>%
  filter(allele =="R89B08")

ggplot(R89B08, mapping = aes(x = totalkernels, y = percentage, color = pollen_handler)) +
  geom_point()
```

```{r}
modR89B08<- lm(percentage ~ pollen_handler + factor(year), R89B08)
summary(modR89B08)

mod2R89B08 <- lm(percentage ~ pollen_handler, R89B08)
summary(mod2R89B08)
```

```{r}
# year is signficant...?
anova(mod2R89B08, modR89B08)
# p-value of .01409
```

```{r}
#R89B08... want to compare J for 2019 and 2020
#p-value: .00138
redR89B08 <- data %>%
  filter(allele =="R89B08") %>%
  filter(pollen_handler == "J") %>%
  filter(year != 2021)
mod2HR89B08 <- lm(percentage ~ factor(year), redR89B08)
summary(mod2HR89B08)
```

```{r}
#p-value: .00219... totalkernels does have effect
mod3R89B08 <- lm(percentage ~ totalkernels, R89B08)
summary(mod3R89B08)
```

```{r}
#signficant diff in year
t.test(percentage ~ factor(year), data = redR89B08)
```

```{r}
# not significant.... interesting

mod4R89B08 <- lm(totalkernels ~ factor(year), redR89B08)
summary(mod4R89B08)

lmod4R89B08 <- lm(log(totalkernels) ~ factor(year), redR89B08)
summary(lmod4R89B08)

```

```{r}
# 2019 is an interesting year...?

mod5R89B08<- lm(percentage ~ factor(year) + totalkernels, redR89B08)
summary(mod5R89B08)

#take this to all years
mod6R89B08 <- lm(percentage ~ factor(year) + totalkernels, R89B08)
summary(mod6R89B08)
```

---R35A03---
there could be a diff in total kernel depending on handler... but there's not really an year (have to check bc of error noted), totalkernel, handler effect on percentage though. 


```{r}
# either compare J and Z or J and M
data %>%
  group_by(allele, pollen_handler, year)%>%
  summarize(count = n(), meankernel = mean(totalkernels), sdkernel = sd(totalkernels), meanp = mean(percentage)) %>%
  arrange(desc(count)) %>%
  filter(allele == "R35A03")
```

```{r}
# visualization of R35A03
R35A03 <- data %>%
  filter(allele =="R35A03")

ggplot(R35A03, mapping = aes(x = totalkernels, y = percentage, color = pollen_handler)) +
  geom_point()
```

```{r}
# confused about outputs for first model for year...?
modR35A03 <- lm(percentage ~ pollen_handler + factor(year), R35A03)
summary(modR35A03)

mod2R35A03 <- lm(percentage ~ pollen_handler, R35A03)
summary(mod2R35A03)
```

```{r}
# refer to earlier comment above
anova(mod2R35A03, modR35A03)
```

```{r}
#R35A03... want to compare Z and J
redR35A03 <- data %>%
  filter(allele =="R35A03") %>%
  filter(pollen_handler == "Z" | pollen_handler == "J")
mod2HR35A03 <- lm(percentage ~ pollen_handler, redR35A03)
summary(mod2HR35A03)
```

```{r}
# p-value: .12
mod3R35A03 <- lm(percentage ~ totalkernels, R35A03)
summary(mod3R35A03)
```

```{r}
# no diff
t.test(percentage ~ pollen_handler, data = redR35A03)
```

```{r}
# both significant ~ .001
mod4R35A03 <- lm(totalkernels ~ pollen_handler, redR35A03)
summary(mod4R35A03)

lmod4R35A03 <- lm(log(totalkernels) ~ pollen_handler, redR35A03)
summary(lmod4R35A03)

# Z is diff from ?... could just be Z...
mod7R35A03 <- lm(totalkernels ~ pollen_handler, R35A03)
summary(mod7R35A03)
```

```{r}
# not significant for both

mod5R35A03 <- lm(percentage ~ pollen_handler + totalkernels, redR35A03)
summary(mod5R35A03)

#take this to all handlers
mod6R35A03 <- lm(percentage ~ pollen_handler + totalkernels, R35A03)
summary(mod6R35A03)
```

---R04A02---
no signficance


```{r}
data %>%
  group_by(allele, pollen_handler, year)%>%
  summarize(count = n(), meankernel = mean(totalkernels), sdkernel = sd(totalkernels), meanp = mean(percentage)) %>%
  arrange(desc(count)) %>%
  filter(allele == "R04A02")
```

```{r}
# visualization of R04A02
R04A02 <- data %>%
  filter(allele =="R04A02")

ggplot(R04A02, mapping = aes(x = totalkernels, y = percentage, color = pollen_handler)) +
  geom_point()
```

```{r}
modR04A02 <- lm(percentage ~ pollen_handler + factor(year), R04A02)
summary(modR04A02)

mod2R04A02 <- lm(percentage ~ pollen_handler, R04A02)
summary(mod2R04A02)
```

```{r}
#no year effect
anova(mod2R04A02, modR04A02)
# p-value of .9187
```

## decided not to compare two handlers

```{r}
# not signficant
mod3R04A02 <- lm(percentage ~ totalkernels, R04A02)
summary(mod3R04A02)
```


```{r}
# no significance
lmod4R04A02 <- lm(log(totalkernels) ~ pollen_handler, R04A02)
summary(lmod4R04A02)

```

```{r}
#take this to all handlers
mod6R04A02 <- lm(percentage ~ pollen_handler + totalkernels, R04A02)
summary(mod6R04A02)
```


--old work---
```{r}
# linear model for one allele (R106E07)
# one handler, but had the most kernels
# this one was not found to be significant in prev analysis
data2 <- data %>%
  filter(allele =="R106E07")
modR106E07 <- lm(percentage ~ totalkernels, data2)
summary(modR106E07)
```
