---
title: "spatial_graphs"
author: "Michelle Bang"
date: '2023-02-21'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The following code is from Nathan. My goal is to produce more graphical representations of different maize using the same framework. 

Bigger goal: create a function that can take in multiple .xml files and produce graphs. 

```{r}
# Function to get a nice df of coordinates 
xml_to_coordinates <- function(input_data_path) {
  library(xml2)
  
  
  #### Importing the data and arranging it into a useful dataframe ####
  input_xml_data <- read_xml(input_data_path)
  
  input_xml_as_list <- as_list(input_xml_data) # Not very elegant, but should work
  
  input_data_type_1 <- input_xml_as_list[[1]][[2]][[2]] # These should be the GFP kernels as long as the person annotating using type "1" for GFP
  input_data_type_1 <- input_data_type_1[-1] # Removing the first sub-list which is just a number string of the data type
  input_data_type_2 <- input_xml_as_list[[1]][[2]][[3]] # These should be non-GFP kernels
  input_data_type_2 <- input_data_type_2[-1]
  
  # Making the dataframes with columns for type, x coordinate, and y coordinate
  type_1_df <- data.frame("type" = rep("GFP", length(input_data_type_1)),
                          "x" = as.numeric(unlist(sapply(input_data_type_1, "[[", 1))), 
                          "y" = as.numeric(unlist(sapply(input_data_type_1, "[[", 2))))
  
  type_2_df <- data.frame("type" = rep("WT", length(input_data_type_2)),
                          "x" = as.numeric(unlist(sapply(input_data_type_2, "[[", 1))), 
                          "y" = as.numeric(unlist(sapply(input_data_type_2, "[[", 2))))
  final_df <- rbind(type_1_df, type_2_df)
  
  return(final_df)
  
}
```

First I will take a look at how the data looks before we do anything. Here we simply have a histogram of the 

```{r}
library(ggplot2);

#testdf <- xml_to_coordinates("~/LabWorkNathan/Data_for_Analysis/all_xml_files/X400x426L-2m2.xml")
testdf <- xml_to_coordinates("/Users/michellebang/st research lab/spatialanalysis/LabWorkNathan-master/Data_for_Analysis/all_xml_files/X400x426L-2m2.xml")

#testdf <- xml_to_coordinates(list[22])

  ggplot(data = testdf, aes(x=type)) +
      geom_bar()  +
      ggtitle("X400x426L-2m2") +
      theme(plot.title = element_text(hjust = 0.5))
```

Next we want to view it in bins. We want at least 15 degrees of freedom. As such, we will mark xmax = kernel with largest x, then ((xmax-xmin)/16) will give us the bins we need.

```{r}
   #calculate amount of length per bin
   num_bins = 30

   xmax = max(testdf$x)
   xmin = min(testdf$x)
   bins = (xmax-xmin)/num_bins
   
   copy_df <- testdf
   
   
   #removes data frame names that would cause issues with the code
   ## not sure what the point of this code is
   ##rm(final_df)
```

```{r}
   #split data into bins by brute 

   #make new data frame for being counted in next section
   final_df = data.frame(bins= seq(xmin,xmax,bins),WT = c(1:(num_bins+1))*0,GFP = c(1:(num_bins+1))*0)
  
  
   #sequences for the for-loops that come next
   sequence = seq(1,nrow(testdf)) 
   sequence2 = seq(1,nrow(final_df)-1)
```

```{r}

   #The brute force part, requires that the final_df is empty or it will double count

     for (val in sequence) {                                             #for each value
        for(i in sequence2) {                                              #for each bin
          
          #if the value is in the bin, but not any bins before it
          if ((copy_df$x[val] >= final_df$bins[i]) && (copy_df$x[val] < final_df$bins[i+1]) ) { 
            if (copy_df$type[val] == "WT") {                            #has to check if the kernel is WT or GFP then increments bin
               final_df$WT[i] = final_df$WT[i] + 1                      
            }               
            else if (copy_df$type[val] == "GFP") {
               final_df$GFP[i] = final_df$GFP[i] + 1
            }
          }
    
        }
     }

     #remove excess row
     final_df = final_df[-c(num_bins+1),]

  
```


```{r}
 #ggplot of the WT with respect to x
 ggplot(data = final_df, aes(x = bins, y = WT)) +
      geom_col()  +
      ggtitle("X400x426L-2m2") +
      ylim(0,15) +
      theme(plot.title = element_text(hjust = 0.5))

```

```{r}
 #ggplot of the GFP with respect to x
 ggplot(data = final_df, aes(x = bins, y = GFP)) +
      geom_col()  +
      ggtitle("X400x426L-2m2") +
      ylim(0,15) +
      theme(plot.title = element_text(hjust = 0.5))
```

----

my contributions:

```{r}
#load file names based on your folder
filenames = list.files("/Users/michellebang/st research lab/spatialanalysis/LabWorkNathan-master/Data_for_Analysis/strange_data", pattern="*.xml", full.names=TRUE)

#extracting path name... will see if we'll use later
pathnames = paste0(dirname(filenames),"/", basename(filenames))
pathnamesdf <- data.frame(pathnames)

print(pathnamesdf[1,])
```
^^ this works omg...

```{r}
#need to figure how how to store multiple files of data... hopefully in one data frame.... currently not working properly...

for (i in 1:27) {
  path <- pathnamesdf[i,]
  name <- paste("testdf", i, sep = "")
  assign(name, xml_to_coordinates(path))
}

#removes faulty xml from the files. After running this you have to run the xml_to_coordinates file loop.

#sequence = rev(seq(1:length(coords)))

#testdf <- xml_to_coordinates(list[22])
```

```{r}
listData = list(testdf1, testdf2, testdf3, testdf4,
                testdf5, testdf6, testdf7, testdf8,
                testdf9, testdf10, testdf11, testdf12, 
                testdf13, testdf14, testdf15, testdf16,
                testdf17, testdf18, testdf19, testdf20,
                testdf21, testdf22, testdf23, testdf24, 
                testdf25, testdf26, testdf27)

titleList <- as.list(basename(filenames))
```

```{r}
#testing to see if the list values are accessible.... it is...
ggplot(data = listData[[1]], aes(x=type)) +
      geom_bar()  +
      ggtitle(titleList[[1]]) +
      theme(plot.title = element_text(hjust = 0.5))
```

```{r}
pdf(file = "/Users/michellebang/st research lab/spatialanalysis/SpatialAnalysisMichelle/strangeInitialPlots.pdf",
    width = 4,
    height = 4,)
for (i in 1:27) {
  print(ggplot(data = listData[[i]], aes(x=type)) +
      geom_bar()  +
      ggtitle(titleList[[i]]) + 
      theme(plot.title = element_text(hjust = 0.5)))
}
dev.off()
```

double checking why X495B-1x402-v.xml, X495B-2x402-v.xml, X495B-2x492-v.xml is showing an empty graph
```{r}
strangedf <- xml_to_coordinates("/Users/michellebang/st research lab/spatialanalysis/LabWorkNathan-master/Data_for_Analysis/strange_data/X495B-5x402-v.xml")

print(strangedf) #empty.....
```
```{r}
#removing the last three elements from the list

listData[c(25,26,27)] = NULL
```

```{r}
#figuring stuff out...
checkcheck = listData[1]
check = checkcheck[[1]]$x

checkk = nrow(checkcheck[[1]]$x)
```

```{r}
#trying out singular case
num_bins = 30
testdf = listData[[1]]

xmax = max(testdf$x)
xmin = min(testdf$x)
bins = (xmax-xmin)/num_bins
   
copy_df <- testdf
  
#split data into bins by brute 

#make new data frame for being counted in next section
final_df = data.frame(bins= seq(xmin,xmax,bins),WT = c(1:(num_bins+1))*0,GFP = c(1:(num_bins+1))*0)
  
  
#sequences for the for-loops that come next
sequence = seq(1,nrow(testdf)) 
sequence2 = seq(1,nrow(final_df)-1)
```

```{r}

   #The brute force part, requires that the final_df is empty or it will double count

     for (val in sequence) {                                             #for each value
        for(i in sequence2) {                                              #for each bin
          
          #if the value is in the bin, but not any bins before it
          if ((copy_df$x[val] >= final_df$bins[i]) && (copy_df$x[val] < final_df$bins[i+1]) ) { 
            if (copy_df$type[val] == "WT") {                            #has to check if the kernel is WT or GFP then increments bin
               final_df$WT[i] = final_df$WT[i] + 1                      
            }               
            else if (copy_df$type[val] == "GFP") {
               final_df$GFP[i] = final_df$GFP[i] + 1
            }
          }
    
        }
     }

     #remove excess row
     final_df = final_df[-c(num_bins+1),]

  
```

```{r}
 #ggplot of the WT with respect to x
 ggplot(data = final_df, aes(x = bins, y = WT)) +
      geom_col()  +
      ggtitle(titleList[[1]]) +
      ylim(0,15) +
      theme(plot.title = element_text(hjust = 0.5))
 
```

```{r}
 #ggplot of the GFP with respect to x
 ggplot(data = final_df, aes(x = bins, y = GFP)) +
      geom_col()  +
      ggtitle(titleList[[1]]) +
      ylim(0,15) +
      theme(plot.title = element_text(hjust = 0.5))
```

```{r}

pdf(file = "/Users/michellebang/st research lab/spatialanalysis/SpatialAnalysisMichelle/strangeGFPWTPlots.pdf",
    width = 4,
    height = 4,)

for (i in 1:24) {
  num_bins = 30
  testdf = listData[[i]]

  xmax = max(testdf$x)
  xmin = min(testdf$x)
  bins = (xmax-xmin)/num_bins
   
  copy_df <- testdf
  
  #split data into bins by brute 

  #make new data frame for being counted in next section
  final_df = data.frame(bins= seq(xmin,xmax,bins),WT = c(1:(num_bins+1))*0,GFP = c(1:(num_bins+1))*0)
  
  
  #sequences for the for-loops that come next
  sequence = seq(1,nrow(testdf)) 
  sequence2 = seq(1,nrow(final_df)-1)
  
  #The brute force part, requires that the final_df is empty or it will     double count

  for (val in sequence) {                                             #for each value
    for(n in sequence2) {                                              #for each bin
          
    #if the value is in the bin, but not any bins before it
      if ((copy_df$x[val] >= final_df$bins[n]) && (copy_df$x[val] < final_df$bins[n+1]) ) { 
        if (copy_df$type[val] == "WT") {                            #has to check if the kernel is WT or GFP then increments bin
          final_df$WT[n] = final_df$WT[n] + 1                      
        }               
          else if (copy_df$type[val] == "GFP") {
            final_df$GFP[n] = final_df$GFP[n] + 1
          }
        }
    
      }
     }

    #remove excess row
  final_df = final_df[-c(num_bins+1),]
  
   #ggplot of the WT with respect to x
print(ggplot(data = final_df, aes(x = bins, y = WT)) +
      geom_col()  +
      ggtitle(titleList[[i]]) +
      ylim(0,15) +
      theme(plot.title = element_text(hjust = 0.5)))
 
 #ggplot of the GFP with respect to x
print(ggplot(data = final_df, aes(x = bins, y = GFP)) +
      geom_col()  +
      ggtitle(titleList[[i]]) +
      ylim(0,15) +
      theme(plot.title = element_text(hjust = 0.5)))
 
}
dev.off()
```